---
title: "Subject Level Summary Table"
output: 
  rmarkdown::html_vignette:
     highlight: haddock
     code_folding: show
     df_print: paged
vignette: >
  %\VignetteIndexEntry{Subject Level Summary Table}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{=html}
<style>.h1, h1{color: #333A73 !important;}</style>
```
```{=html}
<style>.h2, h2{color: #333A73 !important;}</style>
```
```{=html}
<style>.h3, h3{color: #333A73 !important;}</style>
```

# Introduction

This article describes creating reports which can be classified as ADSL Summary Tables using `adsl_summary` function. Example here is created using `ADSL` domain.

**Note:** *All examples assume CDISC ADaM format as input unless otherwise specified.*

# Programming Flow

-   [Read in Data](#readdata)
-   [Preprocess Data](#preprocess)
-   [ADSL Summary Table](#summarisation)
-   [ADSL Table Display](#output)

## Read in Data {#readdata}

```{r message=FALSE, warning=FALSE}
library(carver)
library(dplyr)
library(purrr) # nolint
data(adsl)
```

## Preprocess Data {#preprocess}

Before creating `adsl_summary`, it is required to use the pre-processing function: `mentry` which serves as an essential step in preparing the data for summarisation.

### `mentry()`

The next step in pre-processing involves the usage of `mentry` function to read and process data
with subsets and variables. It takes in the `adsl` as input . The input data gets processed
based on `TRT`(mandatory variable for generating columns in the final table), `BY`(optional variable used to group rows  in the final table) and/or `SUBGROUP`(optional variable used to add subgroups under the columns-TRT in the final table) variables and filtered by subset conditions.

**Note:** *For additional information such as parameters info, refer to the mentry() documentation.*

```{r message=FALSE, warning=FALSE}
mentry_df <- adsl |>
  mentry(
    subset = NA_character_,
    byvar = NA_character_,
    trtvar = "TRT01A",
    trtsort = "TRT01AN",
    subgrpvar = NA_character_,
    trttotalyn = "Y",
    add_grpmiss = "N",
    pop_fil = "SAFFL"
  )
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
dataset_vignette(as.data.frame(mentry_df))
```

## ADSL Summary Table {#summarisation}

This step is used in creating the ADSL summary table using the 'adsl_summary' function. 

The following are the arguments for the table:

-   `datain`: The input data for table generation. The output data from `mentry_df` function.
-   `vars`: Names of `adsl` variables to display in final output which are tilde-separated, Add `"-S"` for numeric variables.
-   `stat_vars`: Statistics to display in table for numeric vars, tilde-separated.
-   `pctdisp`: Denominator to calculate percentages by.
Values: `"TRT", "VAR", "COL", "SUBGRP", "SGRPN", "CAT", "NONE", "NO", "DPTVAR"`.
-  `total_catyn` To return a 'Total' row for categorical analysis in `vars`. Values: `"Y"/"N"`
-  `total_catlabel` Label for total category row. eg- "All"/"Total"
-  `miss_catyn` To include empty/blank values as `miss_catlabel` in categories of `dptvar` variable or not. Values: `"Y"/"N"`
-  `miss_catlabel` Label for missing values
-  `a_subset` Analysis Subset condition; tilde-separated for each variable in `vars`.
-  `denom_subset` Subset condition to be applied to dataset for calculating denominator, tilde-separated for categorical variables within `vars`.

```{r message=FALSE, warning=FALSE}
adsl_sum <- mentry_df |>
  adsl_summary(
    vars = "AGEGR1/AGEGR1N~AGE-S~SEX/SEXN~RACE/RACEN",
    stat_vars = "N~Range~Meansd~Median~IQR",
    pctdisp = "TRT",
    total_catyn = "N",
    total_catlabel = "Total",
    miss_catyn = "N",
    miss_catlabel = "Missing",
    a_subset = "AGE<65~AGE>80~SEX=='F'~NA",
    denom_subset = NA_character_
  )
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
dataset_vignette(as.data.frame(adsl_sum))
```

## ADSL Table Display {#output}

### Prepare data for general table display
This step is used creating the ADSL summary table using 'tbl_processor' function to prepare data for general table display.

The following are the arguments:

- `datain` Input dataframe. The output data from `adsl_summary` function.
- `dptlabel` Tilde-separated labels to set to category variables in data.
If analysis vars (`DPTVAR`) contains `AGEGR1, RACE`; `dptlabel` may be `"Age Group~Race"`.
- `statlabel` Tilde-separated labels corresponding to each Statistic in data.
- `extra_df` Additional dataframe, to merge with `datain`.
- `extra_mergeby` Variables to merge `extra_df` by, if present.
- `dropvars` Variables additional to standard present in input data to be removed from output
- `disp_value_col` Hide/Show value column from the final display.
- `addrowvars` Group Variable(s) to be removed as a column and instead used as row headers in the value column.

### Create flextable output from display templates
This step is used creating the ADSL summary table using 'tbl_display' to create flextable output from display templates.

The following are the arguments:

- `datain` Input dataframe. The output data from `tbl_processor` function.
- `bylabel` Change `BYVAR` names to label, if any.
- `dpthead` String to become name of the column containing categories (`DPTVAL`) in output.
- `font` Font face for text inside table
- `fontsize` Font size for text inside table

### ADSL Table Display Output {#output}
The ADSL Summary Table is generated successfully. Below is the Table generated:

```{r message=FALSE, warning=FALSE}
adsl_sum |>
  display_bign_head(mentry_data = mentry_df) |>
  tbl_processor(
    statlabel = "N~Range~Meansd~Median~IQR",
    dptlabel = "Age Group~NONE~Sex~Race",
    addrowvar = "DPTVAR"
  ) |>
  tbl_display() |>
  flextable::autofit()
```
