---
title: "Adverse Events Tornado Plot"
output: 
   rmarkdown::html_vignette:
     highlight: haddock
     code_folding: show
     df_print: paged
vignette: >
  %\VignetteIndexEntry{Adverse Events Tornado Plot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
h1{color: #333A73 !important;}
h2{color: #333A73 !important;}
h3{color: #333A73 !important;}
</style>

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = "#>"
)
```

# Introduction

This article describes creating the <b>Adverse Events Tornado Plot for Oncology Studies</b>. It can be used for the reports adae_tornado.tot; as well as for CDISC-recommended Tornado Plot. Examples are created using in-built `ADAE` and `ADSL` dataset.

# Programming Flow

-   [Read in Data](#readdata)

-   [Process Data with Subsets and Variables](#processdata)

-   [Tornado Plot Output](#plot)

## Read in Data {#readdata}

To start, load required libraries and the in-built example `adae` and `adsl` dataset.

```{r data}
library(carver)
library(dplyr)
library(rlang)
library(ggplot2)

data("adae")
data("adsl")
```

## Process Data with Subsets and Variables {#processdata}

Before creating Adverse Events `tornado plot`, it's essential to use the `process_tornado_data()`. This function helps prepare the data for visualization, making it a crucial step. With this function, users can customize many parts of the plot to match their specific data and analysis needs.  
While it is recommended that this be used for Oncology data, it can also work with non-Therapeutic Area specific ADaM datasets, which is shown in this example.  
This function block consists mainly `adsl_merge()`, `ae_pre_processor()` and `mentry()` functions to combine ADSL variables, perform adverse-event specific pre-processing and filter records/set grouping variables. It also filters out the maximum severity based on the given inputs. Below, we'll explain the parameters used in `process_tornado_data`:-

- `dataset_adsl`: This parameter is used to pass the variable containing `ADSL` dataset.
 
- `dataset_analysis`: This parameter is used to pass the variable containing analysis dataset. In this example, `ADAE` dataset.
 
- `adsl_subset`: This parameter specifies subset condition(s) for the `ADSL` dataset. It allows the user to filter data before plotting.

- `analysis_subset`: Similar to `adsl_subset`, this parameter specifies subset condition(s) for the analysis dataset.

- `split_by`: This parameter is used for splitting the data by a specific variable. The default value is `NA_character_` which suggests that by default, there's no splitting. If a variable is provided, it would split the data accordingly.
 Example: Region - separate plots would be generated for each region.

- `trtvar`: This parameter represents the treatment variable. It indicates the variable that represents different treatments or groups in the data.

- `trt_left`: This parameter represents the treatment value. It indicates the value that displays on the left hand side of the plot.


- `trt_right`: This parameter represents the treatment value. It indicates the value that displays on the right hand side of the plot.
 
- `trtsort`: This parameter is used for sorting the treatment variable.

- `obs_residual`: This parameter is used to pass a period (numeric) to extend the observation period. If passed as NA, overall study duration is considered for analysis.
eg. if 5, only events occurring upto 5 days past the TRTEDT are considered..

- `fmq_data`: This parameter is required if FMQ terms are to be used as high or low terms, the `FMQ_Consolidated_List` should be passed. Default value is `NULL`.

- `ae_catvar`: This parameter is used for series categorical variable for severity analysis and filter out the maximum severity.

- `denom_subset`: This parameter is a condition which is used to subset data set for calculating denominator. Default value is `NA`.

- `pctdisp`: This parameter is a method that is used to calculate denominator (for %). Valid values: `"TRT"`, `"VAR"`, `"COL"`, `"SUBGRP"`, `"CAT"`, `"NONE"`, `"NO"`, `"DPTVAR"`, `"BYVARxyN"`.
 
- `legendbign`: This parameter determines whether `N` should be displayed in the legend. The default value is `Y` which suggests that by default, `N` is displayed in legend. Default value is `N`.

- `yvar`: This parameter represents the variable to be plotted on the y-axis. The default value is `AESOC`.

NOTE: Default/Example values are provided to give users a starting point. However, these values can be modified as necessary to better suit the specific data and visualization objectives.

The first step for creating `tornado_plot` for Adverse Events is to read the required datasets into the environment. In this example, `ADSL` and `ADAE` are used.
 
Below is an example of calling `process_tornado_data` function for preprocessing the data for tornado plot.

```{r tornado_plot_data, message=FALSE, warning=FALSE}
df <- process_tornado_data(
  dataset_adsl = adsl,
  dataset_analysis = adae,
  adsl_subset = "SAFFL == 'Y'",
  analysis_subset = "TRTEMFL == 'Y'",
  obs_residual = "30",
  fmq_data = NA,
  ae_catvar = "AESEV/AESEVN",
  trtvar = "ARM",
  trt_left = "Xanomeline High Dose",
  trt_right = "Xanomeline Low Dose",
  pctdisp = "TRT",
  denom_subset = NA_character_,
  legendbign = "N",
  yvar = "AESOC"
)
```

The `process_tornado_data` returns `data.frame` which gets stored in `df` above. A pre-processed dataset which is the input for Oncology tornado plot.

<b>df</b>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
dataset_vignette(df)
```

## Tornado Plot {#output}

The `tornado_plot` function is designed to create flipped bar plots from input data, with customizable options. It offers flexibility in plotting data and allows users to tailor the plot aesthetics to their preferences. The parameters used in `tornado_plot` are explained below:

- `datain`: The input data for plotting. The output data from process_tornado_data function.

- `trt_left_label`: This parameter is used to label the `trt_left` in the plot.

- `trt_right_label`: This parameter is used to label the `trt_right` in the plot..

- `bar_width`: Bar width option. This parameter allows customization of the bar width displayed on the plot.

- `axis_opts`: List of options for axis appearance. This parameter allows customization of the appearance of the x-axis and y-axis, including labels, and other options retrieved from plot_axis_opts().

- `legend_opts`: List of options for legend appearance. This parameter allows customization of the legend displayed on the plot, including its  label, position, and direction.

- `series_opts`: Vector of color options with named `ae_catvar` variable for the severity analysis.

- `griddisplay`: Grid display option. This parameter controls whether grid lines are displayed on the plot. Set it to “N” to disable grid lines, and “Y” to enable them.

Below is an example of how `oncology tornado plot` is called using the defined parameters:

```{r, message=FALSE, warning=FALSE}
series_opts <- g_seriescol(df, "red~yellow~blue", "BYVAR1")

plot <- tornado_plot(
  df,
  trt_left_label = "DRUG B",
  trt_right_label = "DRUG C",
  bar_width = 0.5,
  axis_opts = plot_axis_opts(
    xaxis_label = "Primary System Organ Class",
    yaxis_label = "% of Subjects",
    ylinearopts = list(
      breaks = seq(-100, 100, 10),
      labels = c(seq(100, 0, -10), seq(10, 100, 10))
    )
  ),
  legend_opts = list(
    label = "Severity",
    pos = c(0.2, 0.15),
    dir = "vertical"
  ),
  series_opts = series_opts,
  griddisplay = "N"
)
```

```{r, fig.align='left', fig.height=5, fig.width=8, warning=FALSE, results='asis'}
print(plot)
```
